### 2022-10-01 doc/paper.tex

\documentclass[titlepage]{article}

\usepackage{preamble}

\title{
    \textbf{ANALYSIS OF \\
    MACHINE LEARNING \\
    APPLIED TO BOARD GAMES}
}

\author{
    \rule{8cm}{0.3mm} \\[1.3cm] 
    \small \scshape{Adam Amanbaev, Jonathan Hallström, Alvar Edvardsson} \\ 
    \small \scshape{Hugo Åkerfeldt, Romeo Patzer} \\[1cm]
    \small \scshape{Supervisor: Ulf Backlund}
}

\begin{document}

\maketitle

\newpage

\begin{abstract}
In late 2017 DeepMind announced a groundbreaking system in a preprint \cite{alphazero} and the results were astonishing. The system was called AlphaZero and utilized \emph{artificial neural networks} in combination with \emph{heuristic algorithms} in order to teach itself the game chess without any proprietary knowledge. After approximately 9 hours it was able to beat the strongest hand-crafted engines, such as Stockfish and it had learned centuries of human knowledge of chess. In this paper we aim to study the effectiveness of different \emph{neural networks} and \emph{heuristic algorithms} such as the one used in AlphaZero. More precisely, we intend to analyze the efficiency of those networks and algorithms in combination with varying \emph{optimizations, parameters, hyperparameters} and \emph {architectures} applied to the classic games \emph{Connect Four} and \emph{Othello}.

\vskip 0.8cm

\keywords{Machine Learning, Supervised Learning, Reinforcement Learning, Neural Network, Deep Learning}

\end{abstract}

\pagenumbering{roman}

\tableofcontents

\newpage

\section*{\huge Summary of Notation}
\addcontentsline{toc}{section}{Summary of Notation}

\vskip 0.5cm

\epigraph{
I am a forest, and a night of dark trees: but he who is not afraid of my darkness, will find banks full of roses under my cypresses.
\attr{Friedrich Nietzsche, Thus Spoke Zarathustra}
}

\vskip -1cm

\nomenclature{$\upepsilon$}{probability of taking a random action in an $\upepsilon$\hyp greedy policy}
\nomenclature{$\argmin_x f(x)$}{\{$x \mid f(x) = \min_{x'} f(x')$\}}
\nomenclature{$\argmax_x f(x)$}{\{$x \mid f(x) = \max_{x'} f(x')$\}}
\nomenclature{$\leftarrow$}{assignment}
\nomenclature{$\upgamma$}{discount-rate parameter}

\nomenclature{$\textnormal{s, s'}$}{states}

\printnomenclature[3cm]

\clearpage
\pagenumbering{arabic}

\newpage

\section{Introduction}

\subsection{Background}
\subsection{Analysis}

\newpage

\bibliographystyle{plain}
\bibliography{refs}

\end{document}